---
title: "In-class Exercise 3"
author: "kai feng"
date: "Sep 2, 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  freeze: true
---

# Installing R-Packages

```{r}
pacman::p_load(tidyverse, sf, ggstatsplot, tmap, spatstat, raster)
```

# Installing Maptools

eval: false = avoid maptools being download and install repetitively every time the Quarto document been rendered.

```{r}
#| eval: false
install.packages("maptools",
                 repos="https://packagemanager.posit.co/cran/2023-10-13")
```

# Importing Geospatial Data

```{r}
mpsz_sf <- st_read(dsn = "data/MasterPlan2014SubzoneBoundaryWebSHP",                 
               layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
childcare_sf <- st_read("data/child-care-services-geojson.geojson") %>% 
  st_transform(crs=3414)
```

# Creating coastal outline

**st_union() to derive the coastal outline sf tibble data.frame**

```{r}
sg_sf <- mpsz_sf %>% 
  st_union()

plot(sg_sf)
```

# Creating ppp objects from sf data.frame

```{r}
childcare_ppp <- as.ppp(childcare_sf)
plot(childcare_ppp)
```

```{r}
summary(childcare_ppp)
```

# Creating ***owin*** object from sf data.frame

```{r}
sg_owin <- as.owin(sg_sf)
plot(sg_owin)
```

```{r}
summary(sg_owin)
```

# Combining point events object and owin object

```{r}
childcareSG_ppp = childcare_ppp[sg_owin]
plot(childcareSG_ppp)
```

# Kernel Density Estimation of Spatial Point Event

re-scale the unit of measurement to km before performing KDE

```{r}
childcareSG_ppp.km <- rescale.ppp(childcareSG_ppp, 
                                  1000, 
                                  "km")

kde_childcareSG_adaptive <- adaptive.density(
  childcareSG_ppp.km, 
  method="kernel")
plot(kde_childcareSG_adaptive)
```

### convert KDE output into grid object

maptools method

```{r}
par(bg = '#E4D5C9')

gridded_kde_childcareSG_ad <- maptools::as.SpatialGridDataFrame.im(
  kde_childcareSG_adaptive)
spplot(gridded_kde_childcareSG_ad)
```

spatstat.geom method

```{r}
gridded_kde_childcareSG_ad <- as(
  kde_childcareSG_adaptive,
  "SpatialGridDataFrame")
spplot(gridded_kde_childcareSG_ad)
```

# Extracting study area using sf objects

```{r}
pg_owin <- mpsz_sf %>%
  filter(PLN_AREA_N == "PUNGGOL") %>%
  as.owin()

childcare_pg = childcare_ppp[pg_owin]

plot(childcare_pg)  
```
