---
title: "In-class Exercise 4"
author: "kai feng"
date: "Sep 2, 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  freeze: true
---

# Overview

**Spatio-Temporal Point Pattern:** A random collection of points representing time and location of events (eg. disease incidences, species sightings, fires, earthquakes, lightning strikes, tsunamis, volcanic eruptions.)

**Importance:** Increasingly necessary due to growth in geographically and temporally indexed data across various fields.

**Application Example:**

-   Analysis of forest fire events in Kepulauan Bangka Belitung, Indonesia (1 Jan 2023 - 31 Dec 2023).

We want to find out

-   are the locations of forest fire in Kepulauan Bangka Belitung spatial and spatio-temporally independent?

-   if the answer is NO, where and when the observed forest fire locations tend to cluster?

# 1.0 Setup

### 1.1 Installing R-Packages

`sf`: provides functions for importing processing and wrangling geospatial data

`raster`: for handling raster data in R

`spatstat`: for performing Spatial Point Patterns Analysis such as kcross, Lcross, etc.

`sparr`: provides functions to estimate fixed and adaptive kernel-smoothed spatial relative risk surfaces via the density-ratio method and perform subsequent inference. Fixed-bandwidth spatiotemporal density and relative risk estimation is also supported

`tmap`: provides functions to produce cartographic quality thematic maps

`tidyverser`: provide functions to perform common data science tasks including and not limited to data import, data transformation, data wrangling and data visualisation

```{r}
pacman::p_load(sf, raster, spatstat, tmap, tidyverse, sparr)
```

### 1.2 Data Acquisition

We will be using 2 sets of data:

-   **forestfires (CSV file):**

    -   Contains locations of forest fires from MODIS sensor data.

    -   Downloaded from the Fire Information for Resource Management System.

    -   Focus: Only forest fires in Kepulauan Bangka Belitung.

-   **Kepulauan_Bangka_Belitung (ESRI shapefile):**

    -   Shows sub-district boundaries (kelurahan) of Kepulauan Bangka Belitung.

    -   Downloaded from the Indonesia Geospatial Portal.

    -   Focus: Only sub-districts within Kepulauan Bangka Belitung (original data covers all of Indonesia).

### 1.3 Importing And Preparing Study Area/Forest Fire data

::: panel-tabset
## Study Area

### Importing study area

```{r}
kbb <- st_read(dsn = "data/rawdata",        
               layer = "Kepulauan_Bangka_Belitung")
```

To dissolve

```{r}
kbb_sf <- kbb %>% 
  st_union()
```

Drop the Z value in geometry so it become polygon

```{r}
kbb_sf <- kbb_sf %>% 
  st_zm(drop = TRUE, what = "ZM") %>%
  st_transform(crs = 32748)
```

## Forest Fire Data

### Importing Fire Data

```{r}
fire_sf <- read_csv("data/rawdata/forestfires.csv") %>% 
  st_as_sf(coords = c("longitude", "latitude"),
           crs = 4326) %>% 
  st_transform(crs = 32748)
```

Because **ppp** object only accept numerical or character as mark, we need to convert data type of *acq_date* to numeric.

```{r}
fire_sf <- fire_sf %>% 
  mutate(DayOfYear = yday(acq_date)) %>% 
  mutate(Month_num = month(acq_date)) %>% 
  mutate(Month_fac = month(acq_date,
                           label = TRUE,
                           abbr = FALSE))
```
:::

# 2.0 Creating ***owin***

```{r}
kbb_owin <- as.owin(kbb_sf)
kbb_owin
```

```{r}
class(kbb_owin)
```

# 3.0 Visualising the Fire Points

### 3.1 Overall plot

```{r}
tm_shape(kbb_sf)+
  tm_polygons() +
  tm_shape(fire_sf) +
  tm_dots()
```

### 3.2 Visualizing geographic distribution of forest fires by month

```{r}
tm_shape(kbb_sf) +
  tm_polygons() +
  tm_shape(fire_sf) +
  tm_dots(size = 0.1) +
  tm_facets(by="Month_fac",
            free.coords = FALSE, #if change to true, the zoom lvl will be to where the data is
            drop.units = TRUE)
```

### 3.3 Computing STKDE by Month

using `spattemp.density()`

##### 3.3.1 Extracting forest fires by month

\* This will create a new dataset with only the month & geometry data

```{r}
fire_month <- fire_sf %>% 
  select(Month_num)
```

##### 3.3.2 Creating ppp

```{r}
fire_month_ppp <- as.ppp(fire_month)
fire_month_ppp
summary(fire_month_ppp)
```

##### Check for duplicated point events

```{r}
any(duplicated(fire_month_ppp))
```

##### 3.3.3 Including Owin Object

```{r}
fire_month_owin <- fire_month_ppp[kbb_owin]
summary(fire_month_owin)
```

```{r}
plot(fire_month_owin)
```

# 4.0 Computing Spatio-temporal KDE

```{r}
st_kde <- spattemp.density(fire_month_owin)
summary(st_kde)
```

### Plotting the spatio-temporal KDE object

```{r}
tims <- c(7,8,9,10,11,12)
par(mfcol=c(2,3))
for(i in tims){ 
  plot(st_kde, i, 
       override.par=FALSE, 
       fix.range=TRUE, 
       main=paste("KDE at month",i))
}
```

# 5.0 Computing STKDE by Day of Year

### 5.1 Creating ppp object

```{r}
fire_yday_ppp <- fire_sf %>% 
  select(DayOfYear) %>%
  as.ppp()
```

### 5.2 Including Owin object

```{r}
fire_yday_owin <- fire_yday_ppp[kbb_owin]
summary(fire_yday_owin)
```

### 5.3

```{r}
kde_yday <- spattemp.density(
  fire_yday_owin)
summary(kde_yday)
```

```{r}
plot(kde_yday)
```
